---
title: "Permutation test"
author: "Sunny Avry"
date: "June 5th, 2019"
output: html_notebook
--- 

<br>

The goal is to test whether the observed difference between the treatment and the control samples is really significative.See https://www.youtube.com/watch?v=GmvpsJHGCxQ for more details.

```{r}
observed.treatment.sample <- c(19,22,25,26) #the observed scores for the treatment group
observed.control.sample <- c(23,33,40) #the observed scores for the control group
```

```{r}
observed.mean.difference <- mean(observed.treatment.sample) - mean(observed.control.sample)
observed.mean.difference #the mean difference between the two observed groups
```


Is this difference significative or due to chance?


```{r}
whole.sample <- c(observed.treatment.sample,observed.control.sample) #put together the two samples
random.mean.differences <- vector() #this vector will be containing the mean differences of each loop iteration
```


```{r}
for (i in 1:10000) { #compute 10000 times random differences in the whole sample
  
  random.sample <- sample(whole.sample) #randomized the whole sample
  random.treatment.sample <- random.sample[1:4] #selected new random treatment sample
  random.control.sample <- random.sample[5:7] #selected new random control sample
  #for each iteration of the loop, the new mean difference is added to the vector
  random.mean.differences <- c(random.mean.differences, 
                               mean(random.treatment.sample)-
                                 mean(random.control.sample)) 
  
  
}
```


```{r}
hist(random.mean.differences, main = "", xlab = "Sample Means", prob = F, col = "darkred")
abline(v = observed.mean.difference, col = "blue", lwd = 2) #the observed value on the histogram
```


Compute how many times our observed difference is equal or greater than a random difference.

```{r}
p.value.ot <- 1-sum(observed.mean.difference < random.mean.differences) / length(random.mean.differences)
p.value.tt <- 1-sum(abs(observed.mean.difference) > abs(random.mean.differences)) / length(random.mean.differences)

p.value.ot #one-tailed
```
```{r}
p.value.tt #two-tailed
```

In the sample studied, the chance of having the observed difference (-9) or lower is greater than the threshold of 0.05 (one-tailed testing). Similarily, the chance of having a score as extreme as the observed difference is also above the threshold of 0.05 (two-tailed testing).